---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Sweet R tricks
Tips and tricks in R & RStudio, gathered from wherever I see them. Mainly a repo for me to remember cool little R tips I've seen around the place. I hope to update it regularly -- feel free to fork, add your own, and send a PR.

# I/O

### Making Saved Data Smaller
Adding `compress = "xz"` to your `save()` function can make things much smaller. *Very* useful tip from [Ilya Kasnitsky](https://ikashnitsky.github.io/):

```{r}
library(congressbr)
data('senate_nominal_votes')
head(senate_nominal_votes)
save(list = ls(), file = "sen.Rda")
save(list = ls(), file = "sen2.Rda", compress = "xz")

file.info("sen.Rda")$size
file.info("sen2.Rda")$size
```

[Source](https://twitter.com/ikashnitsky/status/973325892956184576)

## Tables

### Totals columns
Nice totals column, from Andrew Heiss:

[source](https://twitter.com/andrewheiss/status/973325552596664321?s=03)

```{r}
library(dplyr)
library(stringr)
library(pander)
mtcars %>% 
  mutate(cars = row.names(.),
         cars = str_extract(cars, "[A-Za-z\\b]*")) %>% 
  count(cars, am) %>% 
  bind_rows(summarise_at(., vars(n), funs(sum)) %>%
              mutate(cars = "**Total**")
  ) %>% 
  pandoc.table()
```

Also, from the comments to the above tweet, by Sam Firke, the author the [janitor](https://github.com/sfirke/janitor) package (I prefer this actually): 
```{r}
library(dplyr)
library(janitor)
mtcars %>% 
  mutate(cars = row.names(.),
         cars = str_extract(cars, "[A-Za-z\\b]*")) %>% 
  count(cars, am) %>% 
  adorn_totals()
```

# ggplot2

You can use curly braces (`{}`) to avail of data wrangling in the middle of your ggplot2 code, as [Alistair](https://stackoverflow.com/users/4497050/alistaire) once explained to me. I can't find the code I was working on at the time, but the idea was to make some slopegraphs. This is a toy/ugly little example:   

```{r message = FALSE, warning = FALSE}
library(dplyr); library(ggplot2)

df <- tibble(
  area = rep(c("Health", "Education"), 6),
  sub_area = rep(c("Staff", "Projects", "Activities"), 4),
  year = c(rep(2016, 6), rep(2017, 6)),
  value = c(15000, 12000, 18000, 24000, 14000, 12000, 13000, 16000, 11000, 
            8000, 15000, 19000)
) %>% arrange(area)

df %>% filter(area == "Health") %>% {
    ggplot(.) +    # add . to specify to insert results here
        geom_line(aes(x = as.factor(year), y = value, 
                      group = sub_area, color = sub_area), size = 2) + 
        geom_point(aes(x = as.factor(year), y = value, 
                       group = sub_area, color = sub_area), size = 2) +
        theme_minimal(base_size = 18) + 
        geom_text(data = dplyr::filter(., 
        year == 2016 & sub_area == "Activities"),    # and here
                  aes(x = as.factor(year), y = value, 
                      color = sub_area, label = sub_area), size = 6, 
                      hjust = 1.2) +
    xlab(NULL) + ylab(NULL) + theme(legend.position = "none")
}
```
[Source](https://stackoverflow.com/questions/44007998/subset-filter-in-dplyr-chain-with-ggplot2)  

Neat little trick from James Goldie -- you can also use `dplyr::case_when()` to highlight certain points on a plot: 

```{r message = FALSE, warning = FALSE}
library(ggplot2); library(ggrepel); library(dplyr)

df <- tibble(
  x = 1:10,
  y = rnorm(10),
  name = c("Apple", "Banana", "Kiwi", "Orange", "Watermelon",
           "Grapes", "Pear", "Canteloupe", "Tomato", "Satsuma")) %>%
  mutate(name_poor = case_when(
    y < 0 ~ name,
    TRUE ~ ""))

ggplot(df, aes(x = x, y = y)) +
  geom_point(size = 3) +
  geom_text_repel(aes(label = name_poor), point.padding = 2)
```
[Source](https://twitter.com/rensa_co/status/976340414016843776?s=08)

## RMarkdown tricks


### R <--> Python
Recently it's been noticed by some [sharp]() observers that you can run Python and R in the same RMarkdown document and that the objects from one are available to the other.

### JavaScript

If you use `results = 'asis'` in the head of your code chunk, i.e. ` ```{r results = 'asis'} `, RMarkdown will keep the result of the chunk 'alive', so to speak, in the document, for you to use. A good example of this is with JavaScript, which you can use to make a nifty [d3]() plot. A few bloggers have noted this, although I think the first was a blogger named [Alice](https://towardsdatascience.com/getting-r-and-d3-js-to-play-nicely-in-r-markdown-270e302a52d3). Here I'll use the example given by [Nick Strayer](http://livefreeordichotomize.com/2017/01/24/custom-javascript-visualizations-in-rmarkdown/). First we save our data to the JSON format:     

```{}

```

We can then use it to draw a d3 graph: 


## RStudio tricks

`ts` plus Shift and Tab gives you a nice data-stamped code section:
```{r}
# Sun Mar 25 14:36:54 2018 ------------------------------
```

[Source](https://community.rstudio.com/t/rstudio-hidden-gems/4974)

